---
title: "BCB5456_R_AssignmentMD"
author: "Shelly Kinney"
date: "2023-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown for BCB 546 R Assignment 2023

# Part 1.1: Data Visualization

First, load up packages tidyverse and tidyr.Then read in files from online for fang_et_al_genotypes.txt and snp_positions.txt.

```{r}
library(tidyverse)

library(tidyr)

snp_pos <- read_tsv("https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2023/main/assignments/UNIX_Assignment/snp_position.txt")

fang_gen <- read_tsv("https://raw.githubusercontent.com/EEOB-BioData/BCB546_Spring2023/main/assignments/UNIX_Assignment/fang_et_al_genotypes.txt", col_names=FALSE)
```

Next is to inspect the files. First, check if the files are lists, then take a head and a tail. Then find the dimensions, the number of rows, and the number of columns. Also check the structure of the files, the names, and the classes.

```{r}
is.list(snp_pos)
is.list(fang_gen)

head(snp_pos)
head(fang_gen)

tail(snp_pos)
tail(fang_gen)

dim(snp_pos)
dim(fang_gen)

nrow(snp_pos)
nrow(fang_gen)

ncol(snp_pos)
ncol(fang_gen)

str(snp_pos)
str(fang_gen)

names(snp_pos)
names(fang_gen)

sapply(snp_pos, class)
sapply(fang_gen, class)
```

For the snp_positions file, the results include:
* is.list: is a list
* head: shows the top 6 rows, first 5 columns and column names
* tail: shows the last 6 rows, first 5 columns and column names
* dim: 983  15
* nrow: 983
* ncol: 15
* str: shows the attributes of each column in the list, such as whether a column contains double or character data
* names: shows the names of each column
* sapply(dataframe, class): shows the class of each column, such as whether a column is numeric or character

For the fang_et_al_genotypes file, the results include:
* is.list: is a list
* head: shows the first 6 rows, first 6 columns and column names
* tail: shows the last 6 rows, first 5 columns and column names
* dim: 2782  986
* nrow: 2782
* ncol: 986
* str: shows the attributes of each column in the list, such as whether a column contains double or character data
* names: shows the names of each column
* sapply(dataframe, class): shows the class of each column, such as whether a column is numeric or character



# Part 1.2: Data Processing

Now the two files must be processed to create 40 files, 20 of maize data and 20 of teosinte data. Each has two sets of separate chromosome files, 10 with increasing position order and ?/? for missing data and 10 with decreasing position order and -/- for missing data.

First, the maize (ZMMIL, ZMMLR, ZMMMR) and teosinte (ZMPBA, ZMPIL, ZMPJA) are to be separated out of the fang_gen file in two separate files (using column 3).

For the snp_pos file, we select out the columns SNP_ID, Chromosome, and Position into a new file.

Next, the maize and teosinte files are to be transposed, and the first row is set to the column names (with the original 1st row being deleted).

The files for maize and teosinte must have the column "Sample_ID" be renamed to "SNP_ID" for merging.

Now we can join the new snp file with both the transposed maize and transposed teosinte files.

Next, using the newly joined maize and teosinte files, we extricate chromosomes 1-10, so that there is one file for each chromosome for both maize and teosinte.

Then sort these files on the position column from smallest to largest.

Next, we copy these files and change ?/? to -/- throughout the file.

Finally, sort these changed files on the position column from largest to smallest.

```{r}
maize_gen <- filter(fang_gen, X3 == "ZMMIL" |X3 == "ZMMLR" | X3 == "ZMMMR"| X3 == "Group")

teo_gen <- filter(fang_gen, X3 =="ZMPBA" | X3 == "ZMPIL" | X3 == "ZMPJA" | X3=="Group")

cut_snp <- select(snp_pos, "SNP_ID", "Chromosome", "Position")

transpose_maize <- as.data.frame(t.data.frame(maize_gen))
colnames(transpose_maize) <- transpose_maize[1,] 
transpose_maize <- transpose_maize[-1,]

##head(transpose_maize)

transpose_teo <- as.data.frame(t.data.frame(teo_gen))
colnames(transpose_teo) <- transpose_teo[1,]
transpose_teo <- transpose_teo[-1,]

##head(transpose_teo)

new_maize <- rename(transpose_maize, SNP_ID = Sample_ID)

new_teo <- rename(transpose_teo, SNP_ID = Sample_ID)

fm <- inner_join(cut_snp, new_maize)

ft <- inner_join(cut_snp, new_teo)

#Clearing environment, keeping full files and original
rm(maize_gen)
rm(teo_gen)
rm(cut_snp)
rm(new_maize)
rm(new_teo)
rm(transpose_maize)
rm(transpose_teo)
#End of environment clearing

mchr1 <- filter(fm, Chromosome == "1")
mchr2 <- filter(fm, Chromosome == "2")
mchr3 <- filter(fm, Chromosome == "3")
mchr4 <- filter(fm, Chromosome == "4")
mchr5 <- filter(fm, Chromosome == "5")
mchr6 <- filter(fm, Chromosome == "6")
mchr7 <- filter(fm, Chromosome == "7")
mchr8 <- filter(fm, Chromosome == "8")
mchr9 <- filter(fm, Chromosome == "9")
mchr10 <- filter(fm, Chromosome == "10")

tchr1 <- filter(ft, Chromosome == "1")
tchr2 <- filter(ft, Chromosome == "2")
tchr3 <- filter(ft, Chromosome == "3")
tchr4 <- filter(ft, Chromosome == "4")
tchr5 <- filter(ft, Chromosome == "5")
tchr6 <- filter(ft, Chromosome == "6")
tchr7 <- filter(ft, Chromosome == "7")
tchr8 <- filter(ft, Chromosome == "8")
tchr9 <- filter(ft, Chromosome == "9")
tchr10 <- filter(ft, Chromosome == "10")

#can make files numeric on position column, but not necessary
#mchr1[,3] <- sapply(mchr1[,3], as.numeric)

imchr1 <- arrange(mchr1, Position)
imchr2 <- arrange(mchr2, Position)
imchr3 <- arrange(mchr3, Position)
imchr4 <- arrange(mchr4, Position)
imchr5 <- arrange(mchr5, Position)
imchr6 <- arrange(mchr6, Position)
imchr7 <- arrange(mchr7, Position)
imchr8 <- arrange(mchr8, Position)
imchr9 <- arrange(mchr9, Position)
imchr10 <- arrange(mchr10, Position)

itchr1 <- arrange(tchr1, Position)
itchr2 <- arrange(tchr2, Position)
itchr3 <- arrange(tchr3, Position)
itchr4 <- arrange(tchr4, Position)
itchr5 <- arrange(tchr5, Position)
itchr6 <- arrange(tchr6, Position)
itchr7 <- arrange(tchr7, Position)
itchr8 <- arrange(tchr8, Position)
itchr9 <- arrange(tchr9, Position)
itchr10 <- arrange(tchr10, Position)

#descending with -/-

dmchr1 <- arrange(mchr1, desc(Position))
dmchr2 <- arrange(mchr2, desc(Position))
dmchr3 <- arrange(mchr3, desc(Position))
dmchr4 <- arrange(mchr4, desc(Position))
dmchr5 <- arrange(mchr5, desc(Position))
dmchr6 <- arrange(mchr6, desc(Position))
dmchr7 <- arrange(mchr7, desc(Position))
dmchr8 <- arrange(mchr8, desc(Position))
dmchr9 <- arrange(mchr9, desc(Position))
dmchr10 <- arrange(mchr10, desc(Position))

dtchr1 <- arrange(tchr1, desc(Position))
dtchr2 <- arrange(tchr2, desc(Position))
dtchr3 <- arrange(tchr3, desc(Position))
dtchr4 <- arrange(tchr4, desc(Position))
dtchr5 <- arrange(tchr5, desc(Position))
dtchr6 <- arrange(tchr6, desc(Position))
dtchr7 <- arrange(tchr7, desc(Position))
dtchr8 <- arrange(tchr8, desc(Position))
dtchr9 <- arrange(tchr9, desc(Position))
dtchr10 <- arrange(tchr10, desc(Position))

dmchr1[dmchr1 == "?/?"] <- "-/-"
dmchr2[dmchr2 == "?/?"] <- "-/-"
dmchr3[dmchr3 == "?/?"] <- "-/-"
dmchr4[dmchr4 == "?/?"] <- "-/-"
dmchr5[dmchr5 == "?/?"] <- "-/-"
dmchr6[dmchr6 == "?/?"] <- "-/-"
dmchr7[dmchr7 == "?/?"] <- "-/-"
dmchr8[dmchr8 == "?/?"] <- "-/-"
dmchr9[dmchr9 == "?/?"] <- "-/-"
dmchr10[dmchr10 == "?/?"] <- "-/-"

dtchr1[dtchr1 == "?/?"] <- "-/-"
dtchr2[dtchr2 == "?/?"] <- "-/-"
dtchr3[dtchr3 == "?/?"] <- "-/-"
dtchr4[dtchr4 == "?/?"] <- "-/-"
dtchr5[dtchr5 == "?/?"] <- "-/-"
dtchr6[dtchr6 == "?/?"] <- "-/-"
dtchr7[dtchr7 == "?/?"] <- "-/-"
dtchr8[dtchr8 == "?/?"] <- "-/-"
dtchr9[dtchr9 == "?/?"] <- "-/-"
dtchr10[dtchr10 == "?/?"] <- "-/-"

#final environment cleanse
rm(mchr1) 
rm(mchr2) 
rm(mchr3)
rm(mchr4) 
rm(mchr5) 
rm(mchr6) 
rm(mchr7) 
rm(mchr8) 
rm(mchr9) 
rm(mchr10) 
rm(tchr1) 
rm(tchr2) 
rm(tchr3) 
rm(tchr4) 
rm(tchr5) 
rm(tchr6) 
rm(tchr7)
rm(tchr8) 
rm(tchr9) 
rm(tchr10)

#commented out below for graphs; remove comment status to run this code
#rm(fang_gen)
#rm(snp_pos)
#rm(fm)
#rm(ft)

#end of environment cleanse


```

All 40 files are now created. 

Naming scheme is as follows: i or d for increasing or decreasing, m or t for maize or teosinte, and chr# for chromosome numbers 1-10.





# Part 2: Data Visualization

A: SNP Density on and across chromosomes, and SNP number in maize vs teosinte.

```{r, echo = FALSE}
##SNP on and across chromosomes, pregraph work

fm <- arrange(fm, Chromosome)
ft <- arrange(ft, Chromosome)

fm[fm == "unknown"] <- "0"
fm[fm == "multiple"] <- "0"
fm[fm == is.na(fm)] <- "0"

ft[ft == "unknown"] <- "0"
ft[ft == "multiple"] <- "0"
ft[ft == is.na(ft)] <- "0"

fm[,3] <- sapply(fm[,3], as.numeric)
ft[,3] <- sapply(ft[,3], as.numeric)

gm <- filter(fm, Position != "0") 
gm[,2] <- sapply(gm[,2], as.numeric)
gm <- arrange(gm, Chromosome)

gt <- filter(ft, Position != "0")
gt[,2] <- sapply(gt[,2], as.numeric)
gt <- arrange(gm, Chromosome)
```

```{r, echo = FALSE}
#Graph for maize
gm <- gm %>% mutate(Position.Binned = cut(Position, 25, labels=FALSE))

ggplot(gm) + geom_bar(mapping = aes(x=Position.Binned)) + facet_wrap(~ Chromosome) + xlab("Area of Chromosome in Maize") + ylab("SNP Density") 
```

```{r,echo=FALSE}
#Graph for teosinte
gt <- gt %>% mutate(Position.Binned = cut(Position, 25, labels=FALSE))

ggplot(gt) + geom_bar(mapping = aes(x=Position.Binned)) + facet_wrap(~ Chromosome) + xlab("Area of Chromosome in Teosinte") + ylab("SNP Density")
```

```{r, echo = FALSE}
##SNP between maize and teosinte

gm <- gm %>% mutate(Group = "Maize", .before=Chromosome)

gt <- gt %>% mutate(Group = "Teosinte", .before=Chromosome)

jgmt <- full_join(gm, gt)

ggplot(jgmt) + geom_bar(mapping = aes(x=Group)) + facet_wrap(~ Chromosome)
```

```{r, echo = FALSE}
ggplot(jgmt) + geom_bar(mapping = aes(x=Position.Binned, fill=Group)) + facet_wrap(~ Chromosome) + xlab("Area of Chromosome") + ylab("SNP Density")
```

B: Missing data and amount of heterozygosity.

```{r, echo = FALSE}

```

C: Similarities between groups from original dataset.

```{r, echo = FALSE}

```

# The END